---
title: "DATA621 Assignment 1"
author: "Irene Jacob"
date: "9/18/2021"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(corrplot)
library("base")
library(MASS)
library(rpart.plot)
library(forecast)
library(GGally)
library(tibble)
library(tidyr)
library(tidyverse)
library(dplyr)
library(reshape2)
library(tidymodels)
```

# Assignment 1

## 1. Data Exploration

```{r}
train <- read.csv("https://raw.githubusercontent.com/irene908/DATA621/main/moneyball-training-data.csv") %>%select(-INDEX)
  
test <- read.csv("https://raw.githubusercontent.com/irene908/DATA621/main/moneyball-evaluation-data.csv") %>%select(-INDEX)
```

```{r}
dim(train)
```

```{r}
summary(train)
```

```{r}
train %>% gather() %>% ggplot(aes(x= value)) + geom_density(fill='light blue') + facet_wrap(~key, scales = 'free')
```


```{r}

train_new <- train %>% gather(key = 'key', value = 'value')

ggplot(train_new,aes(x = key, y = value)) +geom_boxplot()+coord_cartesian(ylim = c(0, 1000))+ theme(axis.text.x=element_text(angle=45, hjust=1))
```

```{r,fig.height = 7, fig.width = 7}
train %>% cor(., use = "complete.obs") %>% corrplot(., type = "upper", diag = FALSE)
```


```{r,fig.height = 7, fig.width = 7}

train %>% gather(key, value, -TARGET_WINS) %>% ggplot(., aes(value, TARGET_WINS)) +  geom_point( color="purple") + geom_smooth(method = "lm", se = FALSE, color = "black") + facet_wrap(~key, scales ="free", ncol = 3) 
```

```{r}

train %>% gather(key, value) %>% filter(is.na(value)) %>% group_by(key) %>% tally() %>% mutate(p = n / nrow(train) * 100) %>% mutate(p = paste0(round(p, ifelse(p < 10, 1, 0)), "%")) %>% arrange(desc(n)) %>% rename(`Variable` = key, `Count` = n, `Percentage` = p) 

```
## 2. Data Preparation

Handling missing data

```{r}
# Drop the BATTING_HBP field
train <- train %>% select(-TEAM_BATTING_HBP)

train_new <- train

train_new$TEAM_PITCHING_SO <- ifelse(train_new$TEAM_PITCHING_SO > 4000, NA, train_new$TEAM_PITCHING_SO)

train_new$TEAM_PITCHING_H <- ifelse(train_new$TEAM_PITCHING_H > 5000, NA, train_new$TEAM_PITCHING_H)

train_new$TEAM_PITCHING_BB <- ifelse(train_new$TEAM_PITCHING_BB > 2000, NA, train_new$TEAM_PITCHING_BB)

train_new$TEAM_FIELDING_E <- ifelse(train_new$TEAM_FIELDING_E > 480, NA, train_new$TEAM_FIELDING_E)

```

```{r}

for(i in 1:ncol(train_new)){
  train_new[is.na(train_new[,i]), i] <- mean(train_new[,i], na.rm = TRUE)
}

train_new <- train_new %>% 
  filter(TARGET_WINS != 0)

```


```{r}

summary(train_new)
```



### Feature Engineering

```{r}

single_Feature <- function(df){ df %>% mutate(TEAM_BATTING_1B = TEAM_BATTING_H - TEAM_BATTING_2B - TEAM_BATTING_3B - TEAM_BATTING_HR) }

train_new <- single_Feature(train_new)
test <- single_Feature(test)
```

### View the final prepared data

```{r}

train_new %>% gather(key, value) %>% ggplot(., aes(value)) + geom_density(fill='blue') + facet_wrap(~key, scales ="free")

```

```{r}

#summary of the prepared train data

summary(train_new)

```




### Transformation

```{r echo=FALSE, fig.width=15, message=FALSE, warning=FALSE}

# created empty data frame to store transformed variables
train_temp <- data.frame(matrix(ncol = 1, nrow = length(train_new$TARGET_WINS)))

# performed boxcox transformation after identifying proper lambda
train_temp$TEAM_BATTING_3B <- train_new$TEAM_BATTING_3B
BATTING_3B_Lambda <- BoxCox.lambda(train_new$TEAM_BATTING_3B)
train_temp$BATTING_3B <- log(train_new$TEAM_BATTING_3B)

# performed boxcox transformation after identifying proper lambda
train_temp$TEAM_BATTING_HR <- train_new$TEAM_BATTING_HR
BATTING_HR_Lambda <- BoxCox.lambda(train_new$TEAM_BATTING_HR)
train_temp$BATTING_HR <- BoxCox(train_new$TEAM_BATTING_HR, BATTING_HR_Lambda)

# performed a log transformation
train_temp$TEAM_PITCHING_BB <- train_new$TEAM_PITCHING_BB
train_temp$PITCHING_BB <- log(train_new$TEAM_PITCHING_BB)

# performed a log transformation
train_temp$TEAM_PITCHING_SO <- train_new$TEAM_PITCHING_SO
train_temp$PITCHING_SO <- log(train_new$TEAM_PITCHING_SO)

# performed an inverse log transformation
train_temp$TEAM_FIELDING_E <- train_new$TEAM_FIELDING_E
train_temp$FIELDING_E <- 1/log(train_new$TEAM_FIELDING_E)

# performed a log transformation
train_temp$TEAM_BASERUN_SB <- train_new$TEAM_BASERUN_SB
train_temp$BASERUN_SB <- log(train_new$TEAM_BASERUN_SB)

train_temp <- train_temp[, 2:13]
train_tmp <- train_temp %>% gather(key = 'key', value = 'value')


ggplot(train_tmp, aes(x=value)) + geom_density() + geom_histogram() + facet_wrap(~key, scales ="free", ncol = 6)

#hist(train_temp)
```

### Finalizing the dataset for model building 


```{r}
# Build clean dataframe with transformation

train_new <- data.frame(cbind(train_new, BATTING_3B = train_temp$BATTING_3B, BATTING_HR = train_temp$BATTING_HR,BASERUN_SB = train_temp$BASERUN_SB, PITCHING_BB = train_temp$PITCHING_BB, PITCHING_SO = train_temp$PITCHING_SO, FIELDING_E = train_temp$FIELDING_E))

is.na(train_new) <- sapply(train_new, is.infinite)

# Impute missing value with the mean

train_new$BATTING_3B[is.na(train_new$BATTING_3B)] <- mean(train_new$BATTING_3B, na.rm = TRUE)
train_new$BASERUN_SB[is.na(train_new$BASERUN_SB)] <- mean(train_new$BASERUN_SB, na.rm = TRUE)
train_new$PITCHING_SO[is.na(train_new$PITCHING_SO)] <- mean(train_new$PITCHING_SO, na.rm = TRUE)
```


## 3. Build models

```{r}
x<-c(1,17,18,19,20,21,22)
train_df <- train_new[,x]
train_new <- train_new[,1:16]

```

### Simple model using the transformed data

selecting a few high correlation variables

```{r}
colnames(train_df)<- c('TARGET_WINS','TEAM_BATTING_3B','TEAM_BATTING_HR', 'TEAM_BASERUN_SB', 'TEAM_PITCHING_BB', 'TEAM_PITCHING_SO', 'TEAM_FIELDING_E')
train_simple <- lm(TARGET_WINS ~ TEAM_BATTING_HR + TEAM_PITCHING_BB + TEAM_FIELDING_E, data = train_df)
summary(train_simple)
par(mfrow = c(2, 2))
plot(train_simple)
```

### Simple model without the transformed data

```{r}

train_simple_t <- lm(TARGET_WINS ~ TEAM_BATTING_HR + TEAM_PITCHING_BB + TEAM_FIELDING_E, data = train_new)
summary(train_simple_t)
par(mfrow = c(2, 2))
plot(train_simple_t)

````

### Full model without the transformed data

```{r}
train_full <- lm(TARGET_WINS ~., data = train_new)
summary(train_full)
par(mfrow = c(2, 2))
plot(train_full)
```



### Polynomial Regression without the transformed data

```{r}

train_poly <- "TARGET_WINS ~ TEAM_BATTING_1B + TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR + TEAM_BATTING_BB + TEAM_BATTING_SO + TEAM_BASERUN_SB + TEAM_BASERUN_CS + TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_PITCHING_BB + TEAM_PITCHING_SO + TEAM_FIELDING_E + TEAM_FIELDING_DP + I(TEAM_BATTING_1B^2)+ I(TEAM_BATTING_2B^2) + I(TEAM_BATTING_3B^2) + I(TEAM_BATTING_HR^2) + I(TEAM_BATTING_BB^2) + I(TEAM_BATTING_SO^2) + I(TEAM_BASERUN_SB^2) + I(TEAM_BASERUN_CS^2) + I(TEAM_PITCHING_H^2) + I(TEAM_PITCHING_HR^2) + I(TEAM_PITCHING_BB^2) + I(TEAM_PITCHING_SO^2) + I(TEAM_FIELDING_E^2) + I(TEAM_FIELDING_DP^2)  + I(TEAM_BATTING_1B^3)+ I(TEAM_BATTING_2B^3) + I(TEAM_BATTING_3B^3) + I(TEAM_BATTING_HR^3) + I(TEAM_BATTING_BB^3) + I(TEAM_BATTING_SO^3) + I(TEAM_BASERUN_SB^3) + I(TEAM_BASERUN_CS^3) + I(TEAM_PITCHING_H^3) + I(TEAM_PITCHING_HR^3) + I(TEAM_PITCHING_BB^3) + I(TEAM_PITCHING_SO^3) + I(TEAM_FIELDING_E^3) + I(TEAM_FIELDING_DP^3)  +I(TEAM_BATTING_1B^4) + I(TEAM_BATTING_2B^4) + I(TEAM_BATTING_3B^4) + I(TEAM_BATTING_HR^4) + I(TEAM_BATTING_BB^4) + I(TEAM_BATTING_SO^4) + I(TEAM_BASERUN_SB^4) + I(TEAM_BASERUN_CS^4) + I(TEAM_PITCHING_H^4) + I(TEAM_PITCHING_HR^4) + I(TEAM_PITCHING_BB^4) + I(TEAM_PITCHING_SO^4) + I(TEAM_FIELDING_E^4) + I(TEAM_FIELDING_DP^4) "
train_poly_lm <- lm(train_poly, train_new)
train_poly_lm_stepback <- MASS::stepAIC(train_poly_lm, direction="backward", trace = F)
train_poly_lm_call <- summary(train_poly_lm_stepback)$call
train_poly_lm_stepback <- lm(train_poly_lm_call[2], train_new)
summary(train_poly_lm_stepback)
par(mfrow = c(2, 2))
plot(train_poly_lm_stepback)
```

### excluding variables with Multicollinearity

```{r}
train_multi <- lm(TARGET_WINS ~.- TEAM_BATTING_SO- TEAM_PITCHING_BB- TEAM_PITCHING_H- TEAM_PITCHING_HR, data = train_new)
summary(train_multi)
par(mfrow = c(2, 2))
plot(train_multi)

```

### Excluding variables having insignificant p values

```{r}

train_p <- lm(TARGET_WINS ~.- TEAM_BATTING_SO - TEAM_PITCHING_BB - TEAM_PITCHING_H - TEAM_PITCHING_HR - TEAM_BASERUN_CS, data = train_new)
summary(train_p)
par(mfrow = c(2, 2))
plot(train_p)

```


## 4. Select Model

```{r}
test$TARGET_WINS <- round(predict(train_poly_lm_stepback, test), 0)
```


```{r}
write.csv(test,"DATA621_Assignment1.csv")
```